SMT-Boosted Security Types for Low-Level MPC

### 2024-10-13 esop25/cpj.tex
checkpoint

\begin{definition}
  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$, we say that
  $X \subseteq S \cup \flips(\prog)$ \emph{interfere with} 
  $M \subseteq V$ iff $\progd(\prog) \not\vdash X * M$.
\end{definition}

\begin{lemma}
  \label{lemma-interference}
  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$,
  for all $x \in S \cup R$ and $M \subseteq V$ exactly one of the following conditions holds:
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\setit{x}$ interferes with $M$.
  \item There exists $X \subseteq S \cup R$ such that $\setit{x} \cup X$ interferes
    with $M$ but neither $X$ nor $\setit{x}$ alone do, in which case we say
    that $\setit{x}$ interferes with $M$ conditionally on $X$.
  \item Neither condition (i) nor (ii) hold.
  \end{enumerate}
\end{lemma}

@@

\begin{definition}
  \label{definition-sound}
  Given $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$
  and $\flips(\prog) = R$, we say that
  \emph{$\Gamma$ is sound for $\prog$} iff for all $M' \subseteq M$
      and $x \in S \cup R$, the following
      conditions hold:
  \begin{enumerate}[\hspace{5mm}i.]
    \item  if $\setit{x}$ interferes with $M'$ then
      there exists $\ty$ with $\leakj{\Gamma}{M'}{\ty}$ and $x \in \ty$.
    \item  if there exists  $X \subseteq S \cup R$ such that
      $\setit{x}$ interferes with $M'$ conditionally on $X$,
      then $\leakj{\Gamma}{M'}{\ty}$ and
      $\leakclose{\Gamma}{\ty \cup X}{\ty'}$ and
      $x \in \ty'$.
  \end{enumerate}
\end{definition}

@@

    \eqj{R_2}{\eqs}{R_2}{\ty_1}
  \end{mathpar}
  These facts imply that $\rx{w}{\cid}$ is used as a one-time-pad to encode
  $\phi$ and introduces interference conditional on $\rx{w}{\cid}$
  \cite{barthe2019probabilistic}. Observing that:
  $$
  \leakclose{\Gamma}{\setit{\cty{\rx{w}{\cid}}{\ty_1}, \rx{w}{\cid}}}{\ty'}
  $$
  The result in this subcase follows by the (second) induction hypothesis. \qed
\end{proof}

>>>>

%\begin{definition}
%  Given $\prog$ and $X_1,X_2 \subseteq \vars{\prog}$, we say that
%  $X_1$ \emph{interferes with} $X_2$
%  iff $\progd(\prog) \not\vdash X_1 * X_2$.
%\end{definition}

\begin{lemma}
  \label{lemma-interference}
  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$ and $\flips(\prog) = R$,
  for all $x \in S \cup R$ and $M \subseteq V$ exactly one of the following conditions holds:
  \begin{enumerate}[\hspace{5mm}i.]
  \item $\notsep{\progd(\prog)}{\setit{x}}{M}$
  \item There exists $X \subseteq S \cup R$ such that
    $\notsep{\progd(\prog)}{\setit{x} \cup X}{M}$ while
    $\sep{\progd(\prog)}{\setit{x}}{M}$ and $\sep{\progd(\prog)}{X}{M}$,
    in which case we say that $x$ encodes $X$ in $M$
  \item Neither condition (i) nor (ii) hold, implying $\sep{\progd(\prog)}{\setit{x}}{M}$.
  \end{enumerate}
\end{lemma}

@@

\begin{definition}
  \label{definition-sound}
  Given $\prog$ with $\iov(\prog) = S \cup M \cup P \cup O$
  and $\flips(\prog) = R$, we say that
  \emph{$\Gamma$ is sound for $\prog$} iff for all $M' \subseteq M$
      and $x \in S \cup R$, the following
      conditions hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item  if $\notsep{\progd(\prog)}{\setit{x}}{M'}$
    there exists $\ty$ with $\leakj{\Gamma}{M'}{\ty}$ and $x \in \ty$.
    \item  if there exists  $X \subseteq S \cup R$ such that
      $\setit{x}$ encodes $X$ in $M'$, 
      then $\leakj{\Gamma}{M'}{\ty}$ and
      $\leakclose{\Gamma}{\ty \cup X}{\ty'}$ and
      $x \in \ty'$.
  \end{enumerate}
\end{definition}

@@

    \eqj{R_2}{\eqs}{R_2}{\ty_1}
  \end{mathpar}
  These facts imply that $\rx{w}{\cid}$ is used as a one-time-pad to encode
  $\vars{\phi}$ in $M'$ 
  \cite{barthe2019probabilistic}. Observing that:
  $$
  \leakclose{\Gamma}{\setit{\cty{\rx{w}{\cid}}{\ty_1}, \rx{w}{\cid}}}{\ty'}
  $$
  The result in this subcase follows by the (second) induction hypothesis. \qed
\end{proof}

### 2024-10-13 esop25/ipj.tex
checkpoint

\begin{lemma}
  If $\prog$ with $\views(\prog) = V$ does not have integrity for some
  $H,C$, then there exists $x \in \houtputs$ with an
  aberration.
\end{lemma}

\begin{lemma}
  Given $\prog$ and $H,C$, if $x$ has an aberration then one of
  the two inductively defined conditions holds:
  \begin{enumerate}[\hspace{5mm}i.]
  \item There exists $\xassign{x}{\be}{\cid} \in \prog$ with $\cid \in C$.
  \item There exists $\xassign{x}{\be}{\cid} \in \prog$ with $\cid \in H$ and
    $x' \in \vars(\be)$ such that $x'$ has an aberration.
  \end{enumerate}
\end{lemma}

\begin{lemma}
  Given pre-processing predicate $\eqspre$ and protocol $\prog$ with
  $\views(\prog) = V$, if 
  $\ipj{\eqs}{\prog}{\Delta}$ is valid
  and $x$ has an aberration for some $H,C$,
  then $\cheatj{\Delta}{H,C}{\seclev}$ and $\seclev(x) = \lolab$.
\end{lemma}

>>>>

\begin{lemma}
  Given protocol $\prog$ with
  $\views(\prog) = V$, if 
  $\ipj{\eqs}{\prog}{\Delta}$ is valid
  for some $\eqs$ and $x \in \cinputs$ has an
  aberration for some $H,C$, then $\cheatj{\Delta}{H,C}{\seclev}$
  and $\seclev(x) = \lolab$.
\end{lemma}

\begin{lemma}
  If $\prog$ with $\iov(\prog) = S \cup V \cup O$ does not have
  integrity for some $H,C$, then there exists $\setit{x} \cup V'
  \subseteq \cinputs$ where $\setit{x}$ has an aberration and
  $\notcondsep{\progd(\prog)}{S_H}{\setit{x} \cup V'}{\houtputs \cup O_H}$.
\end{lemma}

\begin{lemma}
  Given $\prog$ with $\iov(\prog) = S \cup V \cup O$ and
  $\ipj{\eqs}{\prog}{\Delta}$ valid for some $\eqs$.  For all $x \in
  \cinputs$ with an aberration, if
  $\notcondsep{\progd(\prog)}{S_H}{\setit{x} \cup V'}{\houtputs \cup O_H}$
  then there exists $y \in \houtputs \cup O_H$ such that
  $\cheatj{\Delta}{H,C}{\seclev}$ and $\seclev(y) = \lolab$.
\end{lemma}

%\begin{lemma}
%  Given $\prog$ and $H,C$, if $x$ has an aberration then one of
%  the two inductively defined conditions holds:
%  \begin{enumerate}[\hspace{5mm}i.]
%  \item There exists $\xassign{x}{\be}{\cid} \in \prog$ with $\cid \in C$.
%  \item There exists $\xassign{x}{\be}{\cid} \in \prog$ with $\cid \in H$ and
%    $x' \in \vars(\be)$ such that $x'$ has an aberration.
%  \end{enumerate}
%\end{lemma}

### 2024-10-13 esop25/model.tex
checkpoint

\begin{definition}
  Given $\pmf$, the \emph{conditional distribution}
  of $X_1$ given $X_2$ where $X_1 \cup X_2 \subseteq \dom(\pmf)$ and $X_1 \cap X_2 = \varnothing$
  is denoted $\condd{\pmf}{X_1}{X_2}$ and defined as follows:
  $$
  \forall \store \in \mems(X_1 \cup X_2)\ .\ 
  \condd{\pmf}{X_1}{X_2}(\store) \defeq
  \begin{cases}
    0 \text{\ if\ } \margd{\pmf}{X_2}(\store_{X_2}) = 0\\
    \margd{\pmf}{X_1 \cup X_2}(\store) / \margd{\pmf}{X_2}(\store_{X_2})\ \text{\ o.w.}
  \end{cases}
  $$
\end{definition}
We also define some convenient syntactic sugarings. The first will allow us to
compare marginal distributions under different realization conditions
(as in, e.g., Definition \ref{definition-NIMO}), the others are standard.
\begin{definition}
  Given $\pmf$, for all $\store_1 \in \mems(X_1)$ and $\store_2 \in \mems(X_2)$ define:
  \begin{enumerate}
  \item $\condd{\pmf}{X_1}{\store_2}(\store_1) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
  \item $\pmf(\store_1)  \defeq \margd{\pmf}{X_1}(\store_1)$ 
  \item $\pmf(\store_1|\store_2) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
  \end{enumerate}
\end{definition}
As in previous work \cite{darais2019language,barthe2019probabilistic,skalka-near-ppdp24}, 
probabilistic independence, aka separation, is an important concept and we adopt the
following standard notation to express separation:
\begin{definition}
%  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
%  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
  We write $\sep{\pmf}{X_1}{X_2}$ iff for all
    $\store \in \mems(X_1 \cup X_2)$ we have
  $\margd{\pmf}{X_1 \cup X_2}(\store) =
  \pmf(\store_{X_1}) * \pmf(\store_{X_2})$
\end{definition}

>>>>

\begin{definition}
  Given $\pmf$, the \emph{conditional distribution}
  of $X_1$ given $X_2$ where $X_1 \cup X_2 \subseteq \dom(\pmf)$ and $X_1 \cap X_2 = \varnothing$
  is denoted $\condd{\pmf}{X_1}{X_2}$ and defined as follows:
  $$
  \forall \store \in \mems(X_1 \cup X_2)\ .\ 
  \condd{\pmf}{X_1}{X_2}(\store) \defeq
  \begin{cases}
    0 \text{\ if\ } \margd{\pmf}{X_2}(\store_{X_2}) = 0\\
    \margd{\pmf}{X_1 \cup X_2}(\store) / \margd{\pmf}{X_2}(\store_{X_2})\ \text{\ o.w.}
  \end{cases}
  $$
\end{definition}
We also define some convenient syntactic sugarings. The first will allow us to
compare marginal distributions under different realization conditions
(as in, e.g., Definition \ref{definition-NIMO}), the others are standard.
\begin{definition}
  Given $\pmf$, for all $\store_1 \in \mems(X_1)$ and $\store_2 \in \mems(X_2)$ define:
  \begin{enumerate}
  \item $\condd{\pmf}{X_1}{\store_2}(\store_1) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
  \item $\pmf(\store_1)  \defeq \margd{\pmf}{X_1}(\store_1)$ 
  \item $\pmf(\store_1|\store_2) \defeq \condd{\pmf}{X_1}{X_2}(\store_1 \uplus \store_2)$
  \end{enumerate}
\end{definition}
As in previous work \cite{darais2019language,barthe2019probabilistic,skalka-near-ppdp24,li2023lilac}, 
probabilistic independence, aka separation, is an important concept and we adopt the
following standard notation to express conditional and unconditional separation:
\begin{definition}
%  We write $\vc{\pmf}{x}{y}$ iff $\pmf(\{ x \mapsto 0\}\ |\ \{ y \mapsto 0 \}) =
%  \pmf(\{ x \mapsto 1\}\ |\ \{ y \mapsto 1 \}) = 1$.
  We write $\condsep{\pmf}{X_1}{X_2}{X_3}$ iff for all
    $\store \in \mems(X_1 \cup X_2)$ and $\store' \in \mems(X_3)$ we have
  $\margd{\pmf}{X_1 \cup X_2}(\store|\store') =
  \pmf(\store_{X_1}|\store') * \pmf(\store_{X_2}|\store')$. If $X_1$ is empty
  we write $\sep{\pmf}{X_2}{X_3}$.
\end{definition}
